Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-05 paper.tex
fix page 8

\begin{figure}[t]
    \centering
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
    \label{fig:latency_table}
\end{figure}

@@

Surprisingly, the majority of this latency comes from kswapd itself. kswapd goes
through phases of moving pages to the inactive list and reclaiming pages from
the inactive list. This repeats until the amount of free memory increases
beyond a threshold. kswapd does not do any particularly expensive operations
beyond disk I/O. Rather, it executes a bunch of cheaper operations (e.g.
checking if a page was referenced or freeing a page) millions of times in deeply
nested loops. Meanwhile, the workload continues to use up any memory reclaimed
by kswapd. Whenever kswapd is invoked by the page fault handler, it blocks the
faulting process until enough pages are reclaimed.  Moreover, in some cases,
during the \texttt{cont} workload kswapd chooses pages to reclaim much faster
than they can be written to disk.

@@

While large swap files allow greater amount of memory overcommitment, they can
also reduce the degree of disk locality. Thus, large swap files should be
created with disk locality in mind. Some Linux distributions, such as
Ubuntu, have switched to using swap files rather than swap partitions by default
\cite{swap_ubuntu}. Our results suggest that for some workloads this may have a
negative impact on performance if the swap file ends up in non-continuous disk
locations.

% Mark was telling Suhas a TCP joke the other day. He had to keep repeating it slower
% and slower until Suhas got it.

@@

Our results also show that in some cases, kswapd induces more overhead to choose
a good page to reclaim than would be accrued by choosing the wrong page to
reclaim. Most of this overhead seems to be due to the poor asymptotic complexity
of PFRA. Moreover, the swapping subsystem should be aware of disk bandwidth and
latency, since they fundamentally limits how much memory can be reclaimed per
unit time.

Systems with larger memory may conceivably support more memory overcommitment,
so improving the performance of swapping may be highly relavent moving forward.
Additionally, if a system has orders of magnitude more pages, kswapd may spend
significantly more time scanning the active and inactive lists.

@@

In the \texttt{rand} benchmark with hugepages enabled, we observe an overhead
of roughly 0.0002 jiffies per page of memory touched after swapping begins.
Also, kswapd runs for longer periods of time with longer intervals of sleep in
between. This is most likely because the benchmark has some probability of
retouching a page, so free memory is used more slowly, but kswapd also has to
do more work to find pages that have not been recently used.

In the \texttt{cont} benchmark with hugepages disabled, kswapd uses
0.0572 jiffies per page of memory touched.
kswapd runs more frequently and for shorter periods of time because new pages
are touched very quickly, requiring faster reclamation. But because all pages
are only touched once, kswapd has to do very little work to choose pages to
reclaim.

The \texttt{cont} benchmark with hugepages enabled spends
surprisingly little time in kswapd. We suspect that kswapd runs once and choose
a large number of pages to swap out. Subsequently, both the workload and kswapd
wait for I/O. This seems consistent with our previous results, in which some
operations take hundreds of milliseconds to complete, but we have not yet
verified it.

@@

The swapping system in Linux has complicated interactions with workloads and
transparent hugepages. It is not obvious how to optimize a workload for better
swapping behavior. Moreover, we see that under memory pressure, kswapd tends to
run more often during page faults than in its own context, causing very poor
workload performance.

@@

New memory technologies are making it possible to have
terabytes of memory. Recent changes to memory management in Linux demonstrate
focus on improving the scalability of its data structures
\cite{huang2016evolutionary}.

@@

RadixVM tries to mitigate performance issues in highly
concurrent workloads due to serialization of memory management operations on
kernel data structures \cite{radixvm}. This work demonstrates that many parts of
existing memory management schemes are not scalable to larger systems.

>>>>

\begin{figure}[t]
    \centering
    \caption {Tail latency of memory accesses}
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
    \label{fig:latency_table}
\end{figure}

@@

Surprisingly, the majority of this latency comes from kswapd itself. kswapd goes
through phases of maintaining LRU lists and reclaiming pages from
them. This repeats until the amount of free memory increases
beyond a threshold. kswapd does not do any particularly expensive operations
beyond disk I/O. Rather, it executes a bunch of cheaper operations (e.g.
checking if a page was referenced or freeing a page) millions of times in deeply
nested loops. Meanwhile, the workload continues to use up any memory reclaimed
by kswapd. Moreover, in some cases,
during the \texttt{cont} workload kswapd chooses pages to reclaim much faster
than they can be written to disk.

@@

While large swap files allow greater amount of memory overcommitment, they can
also reduce the degree of disk locality. Thus, large swap files should be
created with disk locality in mind. 
% Mark was telling Suhas a TCP joke the other day. He had to keep repeating it slower
% and slower until Suhas got it.

@@

Our results also show that in some cases, kswapd induces more overhead choosing 
a good page to reclaim than would occur by reclaiming the wrong page. Most of this 
overhead seems to be due to the repitition fo cheap operations millions of times.

Systems with larger memory may conceivably support more memory overcommitment,
so improving the performance of swapping may be highly relevant moving forward.
Additionally, in systems with orders of magnitude of more pages, kswapd may spend
significantly more time maintaining LRU lists.

@@

As expected, the \texttt{rand} benchmark with hugepages on incurs the highest
total overhead since it begins swapping very early. For this benchmark we
observe an overhead of roughly 0.0002 jiffies per page memory touched after
swapping begins. Also, kswapd runs for longer periods of time with longer
intervals of sleep in between. This is most likely because the benchmark has
some probability of retouching a page, so free memory is used more slowly, but
kswapd also has to do more work to find pages that are inactive.

In contrast, the \texttt{cont} benchmark with hugepages disabled uses
computational time at roughly 0.0572 jiffies per page of memory touched.
kswapd runs more frequently and for shorter periods of time because new pages
are touched very quickly, requiring faster reclamation, but because all pages
are only touched once, kswapd has to do very little work to choose pages to
reclaim.

Notice also that the \texttt{cont} benchmark with hugepages enabled spends
surprisingly little time in kswapd. We suspect that kswapd runs once and choose
a large number of pages to swap out. Subsequently, both the workload and kswapd
wait for I/O. This seems consistent with our previous results, in which some
operations take on the order of seconds to complete, but we have not yet
verified it.

@@

The swapping system in Linux has complicated interactions with workloads and
transparent hugepages. It is not obvious how to optimize a workload for better
swapping behavior. Moreover, we see that under memory pressure, kswapd tends to
run more often during page faults than in its own context.

@@

The emergence of new technology has in the past often required operating systems
to change. New non-volatile memory technologies are making it possible to have
terabytes of memory. Recent changes to memory management in Linux demonstrate a
lot of focus on improving the scalability of its data structures
\cite{huang2016evolutionary}.

@@

RadixVM tries to overcome performance issues in highly
concurrent workloads due to serialization of memory management operations on
kernel data structures \cite{radixvm}. This work demonstrates that many parts of
existing memory management schemes are not scalable to larger systems.
Similarly, in our work, we examine scalability limitations of memory
management in the Linux kernel.
